Moral Foundations of U.S. Political News Organizations

### 2023-08-16 manuscript_version/revision/manuscript_metapsych_revision.Rmd
added_study3
basic outline of study 3 for erin to add math

>>>>

# Study 3

As a final test of the MFD, we had participants write about explicitly moral topics which was then analyzed with the MFD which supplemented using participant score's on the Moral Foundation Questionnaire. These changes create the best case scenario for finding the theorized differences

# Participants

Participants were recruited in two waves as part of a larger investigation on priming political and religious attitudes (???). Participants were recruited via an online research system (SONA) and were given course credit for their participation. 463 participants were
included in the this study. The study was mostly women (53.9%) and White (76.4%) participants with a mix of minority participants: Black (6.1%), (3.6%), (4.2%), Native American (1.9%), Mixed (2.9%) and Other (4.9%). The average listed age was 19.75 (SD =
200 2.94).

# Materials and Procedure

Data was again collected via Qualtrics. Four fake new stories were presented to participants, which were roughly 400 words each. First, all news stories included a few
sentences describing the use of chemical weapons in the Syrian civil war. The news stories were manipulated with political (Republican v. Democrat) and religious (religious v. not) quotes in a 2 x 2 design. News stories can be found in the online materials. Participants also completed the 30-item version of the MFQ as described in the introduction. In addition to basic demographics (gender, age), participant political orientation was assessed with the same scale described in Experiment 1.

After consenting to participate in the study, participants were randomly shown one of
the four new articles about Syria’s use of chemical weapons. Participants were then asked to
write for 5-10 minutes about their reaction to Syria’s use of chemical weapons and the
needed response from the United States. The second wave of data collection included
different writing prompts designed to capture more of the moral foundation areas in their writing. The following writing prompt was used, “Please write about your attitudes on
abortion (or same-sex marriage or environmentalism) as well as your reason for this stance.”
The three prompts were chosen to create a more varied word set by using topics that should
elicit words from each moral foundations category by soliciting a moral response. In each wave of data collection (Syria prompts, moral prompts) the prompt material was randomized between participants. Participants then completed the MFQ, demographics, and the political
orientation scale.

# Expanding the MFD

In a separate study of 466 university students, participants completed a free association task where we asked them to list all the words that came to mind from *harm/care*, *fairness/cheating*, *loyalty/XX*, *authority/XX*, *purity/XX*. They also completed the 15 item MFQ. Responses to the free association task were cleaned and stemmed. This lead to a new set of moral foundation words: harm-96, fairness-76, ingroup-86, authority-81, and purity-80. These concepts were included in the full dictionary used for Experiment 3.

We additionally created a reduced dictionary that included only concepts correlated with their respective moral foundations scores. Correlations between word frequency and the sum of the MFQ were calculated for each foundation and set of concepts. Words were included in the reduced dictionary if their correlation was two standard deviations away from the mean correlation for that foundation. The reduced dataset included the following number of words for each foundation: harm 4, fairness 3, ingroup 3, authority 4, and purity 2.

```{r word-association}
#need the code for the the above stuff
```

[ERIN, Please do mathy-math here]

# Testing Model Structure

Before testing the theorized differences between liberals and conservatives, we sought to test the validity of the MFD using Multi-Trait, Multi-Method modelling (MTMM) to determine if the MFD is measuring the same latent variables as the MFQ as it should.

```{r mtmm}

# Testing Group Differences

```{r mtmm-groupdiffs}
